The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2019-02-25 Paper.Rmd
typo fixed

We also examined whether the nature of the
tail dependence of an environmental variable across space could influence the 
extinction risk of a metapopulation. We hypothesized that environmental noises
exhibiting greater left-tail dependence across different habitat patches would 
cause higher metapopulation extinction risks because then bad years for the 
component populations would occur simultaneously in many patches,
reducing rescue effects. 
We provided a test of the conceptual validity of this hypothesis using a metapopulation
extension of the Lewinton-Cohen model,
$\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
where $\vec{P}(t)$ is a vector of population densities, 
the $i^{th}$ component $P_i(t)$ representing
the density in the $i^{th}$ habitat patch ($i=1,\ldots,N$) at time $t$. 
The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
$\exp(r+\epsilon_i(t))$. Here $r$ is a growth rate; we used $r=0$. The
$\epsilon_i(t)$ represent environmental noises in each patch. They were standard-normally 
distributed in each patch, were independent through time, and showed the same 
spatial correlations (to within sampling error) for every simulation, 
but were made to exhibit right- or left-tail dependence between patches in
different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
matrix $D$ was a dispersal matrix. For simplicity, we considered habitat patches 
to be arranged in a line, with dispersal occurring either to nearest neighbors only,
or globally, in different simulations. The dispersal 
rate was $d$ (Appendix \ref{SM-ext_risk_model}). After each time step, 
if the population density in a patch was below $1$ (an extinction threshold), it was
set to $0$. We simulated the model 10000 times for each combination of parameters,
starting from initial populations in each patch of X, 
<!--***DAN: Shya, please fill in X-->
and calculated extinction risk after $25$ time steps.

@@

As hypothesized, left-tail dependent environmental fluctuations increased metapopulation
extinction risk for the spatially 
extended Lewinton-Cohen model (Fig. \ref{fig_extrisk_lcohen_local}). 
This was true for both $N=5$ and $N=25$ habitat patches, and for both local 
(Fig. \ref{fig_extrisk_lcohen_local}) and global (Fig. \ref{SM-fig_extrisk_lcohen_global})
dispersal. We also examined the influence of non-N copula structure and tail dependence
on Taylor's law in Appendix \ref{SM-Consequence_TL_simulation}.

\begin{figure}[!h]
\textbf{ \hspace{6.6 cm} (A) \hspace{4.2 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
\caption{Extinction risk for the metapopulation extension of the Lewinton-Cohen model, after 
$25$ time steps, was higher for left-tail-dependent environmental noise. Dispersal was local,
with dispersal rate $d$ (Appendix \ref{SM-ext_risk_model}). Simulations used $N$ patches
for $N=5$ (A) and $N=25$ (B).\label{fig_extrisk_lcohen_local}}
\end{center}
\end{figure}

>>>>

We also examined whether the nature of the
tail dependence of an environmental variable across space could influence the 
extinction risk of a metapopulation. We hypothesized that environmental noises
exhibiting greater left-tail dependence across different habitat patches would 
cause higher metapopulation extinction risks because then bad years for the 
component populations would occur simultaneously in many patches,
reducing rescue effects. 
We provided a test of the conceptual validity of this hypothesis using a metapopulation
extension of the Lewontin-Cohen model,
$\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
where $\vec{P}(t)$ is a vector of population densities, 
the $i^{th}$ component $P_i(t)$ representing
the density in the $i^{th}$ habitat patch ($i=1,\ldots,N$) at time $t$. 
The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
$\exp(r+\epsilon_i(t))$. Here $r$ is a growth rate; we used $r=0$. The
$\epsilon_i(t)$ represent environmental noises in each patch. They were standard-normally 
distributed in each patch, were independent through time, and showed the same 
spatial correlations (to within sampling error) for every simulation, 
but were made to exhibit right- or left-tail dependence between patches in
different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
matrix $D$ was a dispersal matrix. For simplicity, we considered habitat patches 
to be arranged in a line, with dispersal occurring either to nearest neighbors only,
or globally, in different simulations. The dispersal 
rate was $d$ (Appendix \ref{SM-ext_risk_model}). After each time step, 
if the population density in a patch was below $1$ (an extinction threshold), it was
set to $0$. We simulated the model 10000 times for each combination of parameters,
starting from initial populations in each patch of X, 
<!--***DAN: Shya, please fill in X-->
and calculated extinction risk after $25$ time steps.

@@

As hypothesized, left-tail dependent environmental fluctuations increased metapopulation
extinction risk for the spatially 
extended Lewontin-Cohen model (Fig. \ref{fig_extrisk_lcohen_local}). 
This was true for both $N=5$ and $N=25$ habitat patches, and for both local 
(Fig. \ref{fig_extrisk_lcohen_local}) and global (Fig. \ref{SM-fig_extrisk_lcohen_global})
dispersal. We also examined the influence of non-N copula structure and tail dependence
on Taylor's law in Appendix \ref{SM-Consequence_TL_simulation}.

\begin{figure}[!h]
\textbf{ \hspace{6.6 cm} (A) \hspace{4.2 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
\caption{Extinction risk for the metapopulation extension of the Lewontin-Cohen model, after 
$25$ time steps, was higher for left-tail-dependent environmental noise. Dispersal was local,
with dispersal rate $d$ (Appendix \ref{SM-ext_risk_model}). Simulations used $N$ patches
for $N=5$ (A) and $N=25$ (B).\label{fig_extrisk_lcohen_local}}
\end{center}
\end{figure}
